The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-01 paper.tex
minor modifications in the first sections and restructuring of 2.2

The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like
designs of objects and non-human agents can evoke both positive and negative feelings 
of affinity in the human observer / user. Positive feelings reflect the experience of 
emotional engagement with and feelings of empathy for the non-human agent
\citep{cheetham_human_2011}. The positive effect of so-called
\textbf{anthropomorphic form} is exploited in the design of artificial agents
(\eg digital characters, robots) for several reasons. Amongst others,
anthropomorphic design should serve to evoke a positive valence as the Uncanny
Valley hypothesis suggests, which in turn should foster people's acceptance of
the artificial agent, and enable effective (human-like and 
human-social) interaction.  However, at a certain point as the non-human agent
approaches but fails to closely resemble a human, the evoked valence in the
human user suddenly turns into a negative uncanny feeling of what
\cite{mori_uncanny_1970} describes as disgust and revulsion. This, in turn,
might lead to rejection of the non-human but human-like agent.  The Uncanny
Valley hypothesis predicts that with further increasing the human-likeness of
the non-human agent such that it appears so realistic that it is not
distinguishable from a human but perceived to be human, the effect is again a
positive one, reflecting feelings and engagement that are similar as those
evoked by a human-human interaction.  Of course, roboticists and developers try
to avoid ``falling'' into this Uncanny Valley with their human-like designs of
artificial agents. Consequently, there is a considerable research effort to
better understand the boundaries of anthropomorphic design and their effects
on the interaction and experience. To date, findings have been however
inconsistent \citep{cheetham_human_2011}.

Along this line, this article aims at fostering a better understanding of
\textbf{anthropomorphism} (noted \Ant in this paper) by considering the phenomenon
as a whole: robotics researchers often tend to believe that
anthropomorphism only describes a static set of human-like features of a robot
(like shape, speech capabilities, facial expressions, etc.). We refer to these
characteristics as the \emph{anthropomorphic design} of the
robot~\citep{fink_anthropomorphism_2012}. But \emph{anthropomorphism} in general
refers to the larger \emph{social phenomenon} that emerges from the (real or
imagined) \emph{interaction} between a robot and a
user~\citep{persson_anthropomorphism_2000}. According to~\cite{epley_when_2008},
this includes for instance emotional states, motivations, intentions
\emph{ascribed by the user} to the robot. As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with robots.

@@

To support the reflection, it may be useful to consider a few experimental
results. We studied anthropomorphism in several short- and long-term
human-robot interaction studies that we carried out with embodied robots, both
in ecologically valid real-world environments and controlled lab settings. Those
experiments are presented in details in previously published
articles~\citep{fink_anthropomorphic_2012, fink_living_2013, fink2014which,
lemaignan2014dynamics}. We only highlight here a few qualitative results that
underline the complexity of the phenomenon.

\begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-activities.png}
                \caption{Activities related to Roomba, n = 9 households}
                \label{fig:roomba-activities}
        \end{subfigure}%
        \hspace{0.5cm} 
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-perception.pdf}
                \caption{People's perception of Roomba, n = 15 participants}
                \label{fig:roomba-perception}
        \end{subfigure}
    \caption{Usage and perception of the Roomba vacuum-cleaning robot over a 6 months period. Qualitative data suggest that people's activities related to the robot as well as their perception of the robot change over time. Overall, a rather low social engagement of the users with the robot, and few perception of human-likeness in the robot were found.}
    \label{fig:roomba}
\end{figure}

@@

\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.48\columnwidth}
                \centering
                \includegraphics[width=0.75\textwidth]{ranger}
                \caption{Interaction scenario of the \emph{Ranger} domino study: the robot toy box Ranger transports domino tiles between the two children}
                \label{fig:ranger-expe}
        \end{subfigure}%
        \hspace{0.5cm}
        \begin{subfigure}[b]{0.48\columnwidth}
            \includegraphics[width=\textwidth]{ranger-interactions-vs-qualitative-score}
            \caption{Qualitative level of anthropomorphic projections ($y$-axis) versus the actual interactions with the robot ($x$-axis) are negatively correlated.}
            \label{fig:qualitative-score}
        \end{subfigure}
        \caption{Children`s interactions with the Ranger robot toy box, and their anthropomorphic projections onto the robot. Overall, large variations of both children`s behavior and their perception of the robot were found but seemed to be correlated to each other.}
    \label{fig:ranger}
\end{figure}

@@

Anthropomorphism is an interesting phenomenon because it tells us something about 
the relationship between the human who anthropomorphizes, and the non-human agent 
which is anthropomorphized. Though commonly observed, it still appears controversial, at 
the same time, that humans tend to \textit{``see human in the non-
human''} \citep{epley_seeing_2007}. 
To shed light on this phenomenon, first of all, we would like to propose a clarification of terms.
In short, we propose to differentiate between \textbf{anthropomorphic form} to refer to an imitation of human-like form in the \textit{design} of artificial agents, and  \textbf{anthropomorphism} to refer to the social phenomenon of 
people's tendency to \textit{perceive} human-likeness in non human-like agents.
\citep{fink_anthropomorphism_2012} In the following, we might use the terms \texitit{form} and \textit{design} interchangeably.

As anthropomorphism arises in an interaction between a human and a non-human agent
\citep{persson_anthropomorphism_2000}, one can also
take two main perspectives on explaining anthropomorphism \citep{lee_human_2005}: an artifact-centered perspective and a human-centered perspective. 
We now first present an artifact-centered perspective on anthropomorphism, and outline how 
robotics uses \textit{anthropomorphic design} to encourage 
people to interact in a social way with robots.
Then, we present a human-centered 
perspective and shed light on \textit{anthropomorphism as a social phenomenon}. We present 
theories from psychology and cognitive science to explain people's tendency to 
perceive human-like characteristics in non-human agents.
It has to be kept in mind that both perspectives are inseparable; they apply at the 
same time, and we would like to present them here just as two different viewpoints 
on the same phenomenon.
In the end of this section, we synthesize 
the presented backgrounds into an extended understanding of anthropomorphism 
as a multi-factor phenomenon.

@@

An artifact-centered perspective explains anthropomorphism from the ascertainable 
parts of the non-human agent, namely the anthropomorphic form of it.
Using the term \textit{form} to describe the design of an artifact, we refer 
not only to the shape (how it looks like) but to the total 
expression of the artifact \citep{bartneck_shaping_2004}. \cite{disalvo_hug:_2003} 
argues that \textit{form} includes a product's physical shape, materials, and 
behavioral qualities. The behavioral aspect is of special importance for interactive 
systems that show some degree of autonomy (independence).
Several categories of anthropomorphic form can be classified, but there is no 
common agreement on the borders and transitions between these categories: 
anthropomorphic, zoomorphic, caricatured, functional \citep{fong_survey_2003}.
The use of nature-inspired and in particular anthropomorphic (human-like) form in the 
design of non-human things is as old and manifold as people's tendency to anthropomorphize. 
Anthropomorphic (and life-like) forms have a long-standing tradition in design in general, and in the design of daily life things, in particular 
\citep{norman_design_1988,guthrie_bottles_2000}. Interestingly, from its very beginning the image of a robot exists as an outstanding challenge to mankind \citep{replicant_????}. Starting from when the term \texit{robot} first appeared in a play by Karel \^{C}apek in the 1920ies, robots were man-made beings, created as the \textit{``shadow of man''} (and sometimes resembling a human being in appearance).

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize the artifact / 
agent. This perspective assumes that humans directly respond to life-like and social 
cues that the non-human agent or system emits, without thoughtful mental processing, 
by simply applying stereotypes and heuristics to it, and in turn apply 
human-human social schemas and norms to the occurring interactions.
This view is acceptable since humans naturally react to specific 
cues. From early childhood 
on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
or to respond to social cues, such as facial expressions, social gestures or turn-taking. Also \cite{takayama_perspectives_2012} states that humans have a propensity toward particular cues that seem to trigger automatic interpersonal social responses. 
One has to keep in mind that these cues are of various senses and modalities in human-
human interaction, and consequently human-like design also addresses several senses. 
One of the most outstanding aspects is probably the outer appearance of an agent or 
artifact (in particular when it is unknown or unfamiliar). \cite{schmitz_concepts_2011} 
argues that an agent's outer appearance is likely to have an important impact on how 
the object is overall perceived. Consequently, if a robot's outer form resembles a 
human, the robot is likely to be treated similar to a human (\eg eye-contact is 
established or the robot is greeted).

People's natural reaction to anthropomorphic design is exploited in the design of 
robots (in particular, social or also called \textit{sociable} robots 
\citep{breazeal_sociable_2000}) and other things, however, consequences can be both of 
positive and negative nature, which makes anthropomorphic form an interesting subject 
of study.

@@

The underlying central point about the human-like design of robots lies in the 
implications that this design holds, and the suggestions for interactions that
it makes to the human user. 
In other words, the appearance 
and function of a product, whether it is an appliance or a humanoid robot, affects how 
people perceive it, interact with it, and build long-term relationships with 
it \cite{bartneck_shaping_2004}. 
The principle of human-like form is probably most prominent in the design of robots, 
due to several reasons, such as to foster efficient interaction and increase 
acceptance. An anthropomorphic design of a robot encourages people to reason about this 
unfamiliar system \emph{as if} it was a human. To increase this tendency, robots are 
designed with anthropomorphic (or lifelike) forms, or are able to display and use 
human-social cues when interacting with humans.
However, anthropomorphic robot designs have been controversially discussed and received 
remarkable attention since the early 1970ies, related to \cite{mori_uncanny_1970} 
theory of the \textit{Uncanny Valley}, see Figure \ref{fig:uncanny_valley}.

@@

On one hand, developers and designers have discovered that human-like robots may better 
match the affordances that this to humans adapted world brings along. 
Apart from this, it is suggested that increasing the human-likeness of a robot (up to a 
certain threshold) will generally lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970} (see 
Figure \ref{fig:uncanny_valley}).
This further increases anthropomorphism, and as such
appears to be beneficial for human-robot relations, especially with socially 
interactive robots \cite{fong_survey_2003}. Consequently, the domain of (social) 
robotics tries to exploit this by using
anthropomorphic forms in robots.
Humans seem to readily accept and prefer the 
anthropomorphic form of a robot (up to a certain point), when compared to purely mechanical 
robots. A possible explanation for this is that perceiving a system as
human-like (or also pet-like) compared to machine-like changes the 
way how users try to understand and make sense of this system. The reason for this is 
that humans tend to project everyday 
\hl{expectations} of human and animalistic life onto artificial agents that use 
anthropomorphic (or zoomorphic) forms \cite{schmitz_concepts_2011}. 
Both cognitive science 
and social psychology can provide explanations for this process of projection. The basic idea 
is that \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. Studies have further 
shown that perceived similarity between the user and the robot influences the intensity of 
the participants' attributions towards the robot. Thus,the anthropomorphic design of a robot 
impacts the user's perception of the robot, and how far the user is likely to attribute and 
experience qualities such as intelligence, fun, or intentions in the robot. However, this is not a natural spontaneous reaction to anthropomorphic form but already a thoughtful process (investigated as the process of forming a \textit{mental model}), which will be treated in more detail later, in Section \ref{sec:human-centered}).

@@

@@

Literature on anthropomorphism is quite diverse. There exist different understandings of what anthropomorphism is. 
One reason for this might be that anthropomorphism
is a concept that is studied in very different domains that might each
integrate diverse or even contradictory understandings, or simply refer to different things.
However, the conflict is, that despite the fact that there is no commonly accepted definition, the terms
\textit{anthropomorphism}, \textit{anthropomorphic} or \textit{human-like} are
often used as if their meanings were clear and agreed upon
\citep{persson_anthropomorphism_2000}. Duffy \cite{duffy_anthropomorphism_2002} and \cite{epley_when_2008}
argue that these terms might even be misused. For instance, some researchers refer to \textit{"the
robot's level of anthropomorphism"} \cite{bartneck_is_2007}, whereas others
disagree to such a statement because they argue that the system or artifact itself does not \textit{"contain anthropomorphism" per se} but only gives rise to the process of anthropomorphizing in a
given user and situation. Consequently, \cite{persson_anthropomorphism_2000} conclude that
anthropomorphism emerges in the \textit{interaction} between the technology and
the user, and has to be understood as such. 
We will come back to this view later, in the summary of this section (Subsection \ref{sec:multi-factors}).

>>>>

The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like
designs of objects and non-human agents can evoke both positive and negative feelings 
of affinity in the human observer / user. A positive valence reflects the experience of 
emotional engagement with and feelings of empathy for the non-human agent
\citep{cheetham_human_2011}. This in turn may lead to an effective (human-like and 
human-social) interaction with the non-human agent, and increase people's acceptance of
it. 
However, at a certain point as the non-human agent
approaches but fails to closely resemble a human, the evoked valence in the
human user suddenly turns into a negative uncanny feeling of what
\cite{mori_uncanny_1970} describes as disgust and revulsion. This, in turn,
might lead to rejection of the non-human agent. The Uncanny
Valley hypothesis predicts that with further increasing the human-likeness of
the non-human agent such that it appears so realistic that it is not
distinguishable from a human but perceived to be human, the effect is again a
positive one, reflecting feelings and engagement that are similar as those
evoked by a human-human interaction. 
Of course, developers and designers of artificial agents
(\eg virtual characters, robots) try to exploit the positive effects of 
\textbf{anthropomorphic form} and avoid ``falling'' into this Uncanny Valley.
Consequently, there is a considerable research effort to
better understand the boundaries of anthropomorphic design and their effects
on the interaction and experience. To date, findings have been however
inconsistent \citep{cheetham_human_2011}.

A phenomenon that is often studied related to the human-like design of objects and 
artificial agents is \textbf{anthropomorphism}.
Along this line, this article aims at fostering a better understanding of
anthropomorphism (noted \Ant in this paper) in human-robot interaction 
by considering the phenomenon
as a whole. Robotics researchers often tend to believe that
anthropomorphism only describes a static set of human-like features of a robot
(like shape, speech capabilities, facial expressions, etc.). We refer to these
characteristics as the \emph{anthropomorphic design} of the
robot~\citep{fink_anthropomorphism_2012}, and propose that \emph{anthropomorphism} in 
general refers to a \emph{social phenomenon} that emerges from the (real or
imagined) \emph{interaction} between a non-human agent (\eg a robot) and a 
human~\citep{persson_anthropomorphism_2000}. According to~\cite{epley_when_2008},
anthropomorphism includes that the human perceives, for instance, emotional states, motivations, or 
intentions in the non-human agent, and tends to ascribe those qualities to it.
As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with artificial agents such as robots.

@@

To support the reflection on variances and dimensions of anthropomorphism, it may be 
useful to consider a few experimental
results. We studied anthropomorphism in several short- and long-term
human-robot interaction studies that we carried out with embodied robots, both
in ecologically valid real-world environments and controlled lab settings. Those
experiments are presented in details in previously published
articles~\citep{fink_anthropomorphic_2012, fink_living_2013, fink2014which,
lemaignan2014dynamics}. We only highlight here a few qualitative results that
underline the complexity of the phenomenon.

\begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-activities.png}
                \caption{Activities related to Roomba, n = 9 households}
                \label{fig:roomba-activities}
        \end{subfigure}%
        \hspace{0.5cm} 
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-perception.pdf}
                \caption{People's perception of Roomba, n = 15 participants}
                \label{fig:roomba-perception}
        \end{subfigure}
    \caption{Usage and perception of the Roomba vacuum-cleaning robot over a 6 months period. Qualitative data suggest that people's activities related to the robot as well as their perception of the robot change over time.}
    \label{fig:roomba}
\end{figure}

@@

\begin{figure}[h!]
        \centering
        \begin{subfigure}[b]{0.48\columnwidth}
                \centering
                \includegraphics[width=0.75\textwidth]{ranger}
                \caption{Interaction scenario of the \emph{Ranger} domino study: the robot toy box Ranger transports domino tiles between the two children}
                \label{fig:ranger-expe}
        \end{subfigure}%
        \hspace{0.5cm}
        \begin{subfigure}[b]{0.48\columnwidth}
            \includegraphics[width=\textwidth]{ranger-interactions-vs-qualitative-score}
            \caption{We found a significant negative correlation ($p=0.048$) between the
    number of interactions and the qualitative anthropomorphic projections. n = 13 groups}
            \label{fig:qualitative-score}
        \end{subfigure}
        \caption{Children`s interaction with Ranger, and their anthropomorphic projections onto the robot. Overall, large variations of both children`s behavior and their perception of the robot were found but seemed to be correlated to each other.}
    \label{fig:ranger}
\end{figure}

@@

Anthropomorphism is an interesting phenomenon because it tells us something about 
the relationship between the human who anthropomorphizes, and the non-human agent 
which is anthropomorphized. Though commonly observed, it still appears controversial, at 
the same time, that humans tend to \textit{``see human in the non-
human''} \citep{epley_seeing_2007}. 
To shed light on this phenomenon, first of all, we would like to propose a clarification of terms.
In short, we propose to differentiate between \textbf{anthropomorphic form} to refer to an imitation of human-like form in the \textit{design} of artificial agents, and  \textbf{anthropomorphism} to refer to the social phenomenon of 
people's tendency to \textit{perceive} human-likeness in non human-like agents.
\citep{fink_anthropomorphism_2012} In the following, we might use the terms \texit{form} and \textit{design} interchangeably.

As anthropomorphism arises in an interaction between a human and a non-human agent
\citep{persson_anthropomorphism_2000}, one can also
take two main perspectives on explaining anthropomorphism \citep{lee_human_2005}: an artifact-centered perspective and a human-centered perspective. 
We now first present an artifact-centered perspective, and outline how 
robotics uses \textit{anthropomorphic design} to encourage 
people to interact in a social, human-like way with robots.
Then, we present a human-centered 
perspective and shed light on \textit{anthropomorphism as a social phenomenon}. We present 
theories from psychology and cognitive science to explain people's tendency to 
perceive human-like characteristics in non-human agents.
It has to be kept in mind that both perspectives are inseparable; they apply at the 
same time, and we would like to present them here just as two different viewpoints 
on the same phenomenon.
In the end of this section, we synthesize 
the presented backgrounds into an extended understanding of anthropomorphism 
as a multi-factor phenomenon.

@@

An artifact-centered perspective explains anthropomorphism from the ascertainable 
parts of the non-human agent, namely the anthropomorphic form of it.
Using the term \textit{form} to describe the design of an artifact, we refer 
not only to the shape (how it looks like) but to the total 
expression of the artifact \citep{bartneck_shaping_2004}. \cite{disalvo_hug:_2003} 
argues that \textit{form} includes a product's physical shape, materials, and 
behavioral qualities. The behavioral aspect is of special importance for interactive 
systems that show some degree of autonomy (independence).
Several categories of anthropomorphic form can be classified, but there is no 
common agreement on the borders and transitions between these categories: 
anthropomorphic, zoomorphic, caricatured, functional \citep{fong_survey_2003}.
The use of nature-inspired and in particular anthropomorphic (human-like) form in the 
design of non-human things is as old and manifold as people's tendency to anthropomorphize. 
Anthropomorphic (and life-like) forms have a long-standing tradition in design in general, and in the design of daily life things, in particular 
\citep{norman_design_1988,guthrie_bottles_2000}. Interestingly, from its very beginning robots were resembling a human being in their appearance. Starting from when the term \texit{robot} first appeared in a play by Karel \^{C}apek in the 1920ies, robots were man-made beings, created as the \textit{``shadow of man''} \citep{replicant_????}.

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize it. 
This perspective assumes that humans directly (mindlessly) respond to life-like and social 
cues that the non-human agent or system emits \citep{nass_machines_2000}. Without thoughtful mental processing, 
humans tend to simply apply stereotypes and heuristics to the non-human agent, and in turn apply human-human social schemas and norms to the occurring interactions.
\cite{takayama_perspectives_2012} propose to refer to this immediate (sometimes visceral) sense in a situation as \textit{in-the-moment}, compared to a more distanced cogitation and consideration which she calls \textit{reflective}. (As this distinction imply different cognitive processes, we will come back to it later, in the human-centered perspective, Section \ref{sec:human-centered}).

The view of people mindlessly reacting in a human-social manner to anthropomorphic forms
is based on the fact that humans naturally react to specific 
cues. From early childhood 
on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
or to respond to social cues, such as facial expressions, social gestures or turn-taking. Humans have a propensity toward particular cues that seem to trigger automatic interpersonal social responses \citep{takayama_perspectives_2012}. 
One has to keep in mind that in 
human-human interaction these cues are of various senses and modalities, 
and consequently human-like design is also realized addressing several senses. 
One of the most outstanding aspects is probably the outer appearance of an agent or 
artifact (in particular when it is unknown or unfamiliar). \cite{schmitz_concepts_2011} 
argues that an agent's outer appearance is likely to have an important impact on how 
the object is overall perceived. Consequently, if a robot's outer form resembles a 
human, the robot is likely to be treated similar to a human. Besides the physical shape of a robot, other more subtle anthropomorphic forms can be found in robots, including human-social behavioral cues (\eg eye-gaze), social gestures (\eg waving), or verbal politeness (\eg greeting, apologizing).

People's natural reaction to these and other anthropomorphic design cues of
robots (in particular, social or also called \textit{sociable} robots 
\citep{breazeal_sociable_2000}) are extensively studied. This research is of particular interest in human-robot interaction as the consequences of anthropomorphic design can be both of positive and negative nature, and it is yet not clear what exactly it is that makes the difference.

@@

The underlying central point about the human-like design of robots lies in the 
implications that this design holds, and the suggestions for interactions that
it makes to the human user. 
In other words, the appearance 
and function of a product, whether it is an appliance or a humanoid robot, affects how 
people perceive it, interact with it, and build long-term relationships with 
it \citep{bartneck_shaping_2004}. 
The principle of human-like form is probably most prominent in the design of robots, 
due to several reasons, such as to foster efficient interaction and increase 
acceptance. An anthropomorphic design of a robot encourages people to reason about this 
unfamiliar system \emph{as if} it was a human. To increase this tendency, robots are 
designed with anthropomorphic (or lifelike) forms, or are able to display and use 
human-social cues when interacting with humans.
However, anthropomorphic robot designs have been controversially discussed and received 
remarkable attention since the early 1970ies, related to \cite{mori_uncanny_1970} 
theory of the \textit{Uncanny Valley}, see Figure \ref{fig:uncanny_valley}.

@@

On one hand, developers and designers have discovered that human-like robots may better 
match the affordances that this to humans adapted world brings along. 
Apart from this, it is suggested that an increased human-likeness of a robot (up to a 
certain threshold) will generally lead to 
more perceived \textit{affinity} from the human side \citep{mori_uncanny_1970} (see 
Figure \ref{fig:uncanny_valley}).
This further increases anthropomorphism, and as such
appears to be beneficial for human-robot relations, especially with socially 
interactive robots \citep{fong_survey_2003}. Consequently, the domain of (social) 
robotics tries to exploit this by using
anthropomorphic forms in robots.
Humans seem to readily accept and prefer the 
anthropomorphic form of a robot (up to a certain point), when compared to purely mechanical 
robots. A possible explanation for this is that perceiving a system as
human-like (or also pet-like) compared to machine-like changes the 
way how users try to understand and make sense of this system. The reason for this is 
that humans tend to project everyday 
\hl{expectations} of human and animalistic life onto artificial agents that use 
anthropomorphic (or zoomorphic) forms \citep{schmitz_concepts_2011}. 
Both cognitive science 
and social psychology can provide explanations for this process of projection. The basic idea 
is that \textit{"the more similar in 
appearance, the more people are likely to use themselves as a source of induction 
and anthropomorphize these nonhuman agents"} \citep{epley_seeing_2007}. Studies have further 
shown that perceived similarity between the user and the robot influences the intensity of 
the participants' attributions towards the robot. Thus, the anthropomorphic design of a robot 
impacts the user's perception of the robot, and how far the user is likely to attribute and 
experience qualities such as intelligence, fun, or intentions in the robot. 
These experiences mark the transition between an immediate \textit{in-the-moment} and more thoughtful \textit{reflective} \citep{takayama_perspectives_2012} perception of, and interaction with the robot. As such, a more human-centered perspective needs to be applied, and we will come back to this in Section \ref{sec:human-centered}.

@@

But why do people anthropomorphize? First and foremost, it needs to be mentioned that anthropomorphism is a fairly complex phenomenon. Paradoxically, there is a huge and diverse body of literature on it but nevertheless, it seems that anthropomorphism is still not very well understood (especially in a field as young as HRI).

One common reason for anthropomorphism is that by perceiving some human-like characteristics in the non-human agent, this agent is made more \emph{graspable}, \emph{understandable}, \emph{predictable}, and one can feel \emph{more empathetic} toward it, and possibly one can then anticipate the agent's behavior. These are central motivations when trying to explain anthropomorphism, and we will come back to them in more detail later on. 
It is described that for the just mentioned reasons, people commonly anthropomorphize their pets, by ascribing mental states, intentions or feelings to them\footnote{Interestingly, pet-ownership seems to have a significant impact on human-robot interaction: several studies suggest that pet-owners are more likely to anthropomorphize technologies and robots than non pet-owners} \cite{eddy_attribution_1993}. But how can these motivations behind anthropomorphism be explained from a human-centered point of view?

%Anthropomorphism as a social phenomenon arises in an interaction \citep{persson_anthropomorphism_2000}, and thus includes at least two interaction partners (the human and the real or imagined non-human agent). Therefore, one can take two main perspectives on explaining it \citep{lee_human_2005}: a human-centered, and an artifact-centered. We briefly outline both views in the following. It has to be kept in mind that both perspectives are inseparable; they apply at the same time, and we would like to present them here just as two different viewpoints on the same kind of phenomenon.

This perspective builds on research from developmental and social psychology as well as cognitive sciences.  In the following, we present what we found in reviewing this literature: (1) a line of research in developmental psychology that studies the attribution of animacy; (2) a line of research in more cognitive sciences that studies the cognitive processes that underlie anthropomorphism, and (3) a theory that explains anthropomorphism based on psychological factors~\citep{epley_seeing_2007}. All three branches apply a human-centered perspective on anthropomorphism.

\paragraph{Developmental psychology research on the attribution of animacy and intention\\}

The phenomenon of attributing animacy and intentions to nonliving objects such as simple shapes has been intensively studied in developmental psychology. As mentioned earlier, Piaget found that children tend to ascribe life to the nonliving. They also tend to interpret physical phenomena in terms of intention on the part of the non-human subject. For instance children are likely to argue that \textit{``the sun is hot because it wants to make people warm''} \citep{leeds_childrens_1992}. In this example it is in fact questionable whether children really anthropomorphize the sun (believing that it has an intention) or whether they are using anthropomorphism in a metaphoric sense (because their conceptions of the world and of the living are not yet fully formed). Later however, two famous experiments by \cite{heider_experimental_1944} and \cite{michotte_perception_1963} showed that also adult participants attribute animacy and intention to simple things and shapes based on motion.\footnote{For more details, the reader may refer to the ``attribution theory''.} In both experiments, participants viewed animations of simple shapes, such as circles or triangles. 
%(see Figure \ref{fig:animacy_attribution})
Asked to describe what they observed, most people developed elaborate stories, attributing motivations, emotions and relationships between the objects.

Summing it up, as mentioned earlier, it seems that there is a natural, relatively spontaneous reaction that accounts for people's tendency to attribute animacy and intention to moving shapes, even when there is no similarity of the shape to a human. This tendency to animize / anthropomorphize is already developed (and even more distinct) in infants and children. Thus besides the anthropomorphic cues emitted by the design of the artifact, also developmental stages, and people's natural reaction play a role how far we interpret movements / motions (as one of the most basic "lifelike" cues), for instance.

\paragraph{Cognitive science research, the mental model theory (Theory of Mind)
and neural correlates of anthropomorphism\\}

Anthropomorphism and similar phenomenons can also be understood as a thoughtful process of induction whereby \textit{``people reason about an unknown stimulus based on a better-known representation of a related stimulus"}~\citep{epley_when_2008}. In the case of anthropomorphism this means a person's reasoning about a non-human agent based on the representation of the self or other humans. A central construct of this perspective is the so-called \textit{mental model} that a person has (and builds) of the agents she is reasoning about and interacting with. The specific mental model we have about an artifact / agent basically explains (in our individual way to ourselves) how this artifact / agent works the way it does.
There are two central aspects here, when looking at the quote from \citep{epley_when_2008} above.\\
\textit{(1)} One assumption is that the user is not familiar with the other agent. This is basically true for all other agents (and entities) but oneself because we can only be our self. This view is also consistent with the \emph{familiarity thesis}~\citep{hegel_understanding_2008} which claims that we understand the world based upon a mental model of the world that we are most familiar with (our self). Thus, we can best use our self as source of induction when reasoning about other agents. However, in case we would like to understand the behavior of a spider, we would probably not use our mental model of a human (or our self) simply because we do not have many things in common with a spider. And this is the second central aspect here.\\
\textit{(2)} One other assumption is based on the anthropomorphic design of the other agent (see Section \ref{sec:anthropomorphic-design}): if the agent behaves (appears) much like a human being (\eg a robot that emits a human voice), people's mental model of the agent's behavior is likely to approach their mental model of humans (though the model may differ in some important aspects).

This cognitive viewpoint on anthropomorphism is important because people's estimation of an agent's knowledge model and its capabilities affects the way they relate to it. This holds important implications for the resulting interaction with the system, the user experience, and the acceptance. Previous research examined the validity of the mental model concept with various kinds of robots \citep{schmitz_concepts_2011,kiesler_mental_2002}. Findings suggest that people tend to hold richer mental models of anthropomorphic robots in contrast to mechanic ones \citep{kiesler_mental_2002}.\footnote{\hl{Later in this article, we may however come to question how far this result is based on expectations from the human user side, which are likely to be refined after continued interaction, and after the user gets acquainted with the robot's behavior.}} 
A similar finding is described in \cite{hegel_understanding_2008} and
\citet{krach_can_2008}. In a user study using functional magnetic resonnance
imagery (fMRI), the authors found that participants
implicitly attributed human-like qualities, such as mental states, to their
non-human interaction partner. Indeed, from an early age humans develop a
tendency to explain one's own and others'actions in terms of beliefs, desires
and goals,  called \textit{theory of mind}. The theory of mind allows the implicit attribution of
intentions and other mental states to others \citep{premack1978does, leslie_pretense_1987, Frith2003}.
\citet{hegel_understanding_2008} and \citet{krach_can_2008} finding was evident at the behavioral level
as well as on the neuro-physiological level: the more the human-likeness of the
artefactual partner increases, the more brain areas associated with theory of
mind  get activated
\citep{krach_can_2008}. \footnote{\hl{Ha! This says that anthropomorphism can also occur on a behavioral level, yes! :) so we should also measure anthropomorphism in the behavior.}} 
Implication of human brain area involved in the inference of others` mental
states has also been shown in response to viewing
non-anthropomorphic/non-humanoid robotic gadgets whose behavior
has been described as unpredictable -- but not in response to those whose
behavior was described as predictable -- a finding also found at the behavioral
level \citep{Waytz2010}.
Interestingly, the grey matter volume of a brain area related to theory of mind
has been correlated to individual's score of anthropomorphism
\citep{cullen2013individual}.

At lower, more automatic, level, robots have been shown to elicit resonance behaviors in the
human brain. Resonance behaviors \citep{Rizzolatti1999} are mechanisms by which the brain areas involved in the internal representations of an action, an emotion or a sensation are equally recruited during the perception of another individual performing the same action or experiencing the same emotion or sensation.  
The neurons showing resonant properties have been called mirror neurons and
have been found in a wide range of modalities (motor, emotional etc..)
\fixme{addref}. Seeing a robotic arm reaching for an object activates the motor mirror neurons the same way as for seeing a human arm reaching for the same object
\citep{Gazzola2007, oberman_eeg_2007}. However, a study of emotion perception
on a robotic face found reduced activity in emotional brain
area known to have mirror properties in response to the robotic face compared to
a human face \citep{Chaminade2010}. Nevertheless, these authors showed that when
the participants were explicitly instructed to pay attention to the robot's
emotional expression, motor mirror neurons get activated.
Interestingly, using electroencephalography,  it has also been shown that
emotional behavior elicited by a robot face influences the speed of human
responses as well as early brain processes the same way as emotional expression
of a human face did \citep{Dubal2010}.

In sum, this branch of research suggests that anthropomorphism implies that
people thoughtfully develop a mental model of agents in their environment and
that they make inferences about these agents based on what is familiar to them
-- themselves, humans and human behavior. This understanding builds on the
theory of mind,  a
person's ability to attribute mental states to oneself and others. 
%A theory of mind for other agents enables us to attribute intentionality to those agents \citep{leslie_pretense_1987,admoni_multi-category_2012}. 
We briefly reviewed  neuro-physiological evidence of a link between the tendency
to thoughtfully anthropomorphize and the engagement of brain processes involved
in the attribution of mental states to other humans. The human-likeness quality
of the artefactual agent seems to play a role in the involvement of theory of
mind but also the unpredictable feature of the agent behavior alone seems to be
sufficient. Low-level mechanisms of the brain that allow humans to map others`
motor behaviors into their own repertoire are equally elicited when perceiving a
robot's action. It is not yet clear whether humans are processing emotional
signals from a robot the same way as they do for human beings. 
%were recently demonstrated at the brain level~\citep{cullen2013individual}.

\paragraph{Social psychology research: a 3-factor theory of anthropomorphism\\}
\label{sec:psychological-factors}

\begin{enumerate}
	\item ~anthropocentric knowledge is accessible and applicable to the artifact (\textit{elicited agent knowledge}),
	\item ~they are motivated to explain and understand the behavior of other agents (\textit{effectance motivation}), and
	\item ~they have the desire for social contact and affiliation (\textit{social motivation}).
\end{enumerate}

@@

We have seen that literature on anthropomorphism is quite diverse. We would like to point out that there exist different understandings of what anthropomorphism is. 
One reason for this might be that anthropomorphism
is a concept that is studied in very different domains that might each
integrate diverse or even contradictory understandings, or simply refer to different things.
However, the conflict is, that despite the fact that there is no commonly accepted definition, the terms
\textit{anthropomorphism}, \textit{anthropomorphic} or \textit{human-like} are
often used as if their meanings were clear and agreed upon
\citep{persson_anthropomorphism_2000}. Duffy \cite{duffy_anthropomorphism_2002} and \cite{epley_when_2008}
argue that these terms might even be misused. For instance, some researchers refer to \textit{"the
robot's level of anthropomorphism"} \cite{bartneck_is_2007}, whereas others
disagree to such a statement because they argue that the system or artifact itself does not \textit{"contain anthropomorphism" per se} but only gives rise to the process of anthropomorphizing in a
given user and situation. Consequently, \cite{persson_anthropomorphism_2000} conclude that
anthropomorphism emerges in the \textit{interaction} between the technology and
the user, and has to be understood as such. 
We will come back to this view later, in the summary of this section (Subsection \ref{sec:multi-factors}).
